Ages of Planets and Stars Observed by Kepler

### 2024-04-15 paper/ms.tex
update ms

\author{Authors to include (order after LGB not assigned meaning;
currently alphabetical)}

% Early experiments for Kepler rotation samples
\author[0000-0001-7967-1795]{Elsa~K.~Palumbo}
\affiliation{\caltech}

% ?
%\author{Lynne~A.~Hillenbrand}
%\affiliation{\caltech}

% HIRES observations and reductions
\author[0000-0002-0531-1073]{H. Isaacson}
\affiliation{\berkeley}
% HIRES observations, reductions, JUMP infrastructure
\author[0000-0001-8638-0320]{A. W. Howard}
\affiliation{\caltech}
%
\author{HIRES observers to be named}
\affiliation{Various}

@@

A limitation of the ``cluster-first'' approach is that only
$\approx$1\% of stars within the nearest few hundred parsecs can
currently be associated with their birth cluster
\citep[e.g.][]{Zari_2018,CantatGaudin_2020,Kounkel_2020,Kerr_2021}.
This means that there are many ``young'' (0.05-1\,Gyr) stars in the
field that are omitted by cluster-only searches.
Of order 40 transiting planets are currently known to be younger than
1\,Gyr, requiring an age precision $t/\sigma_t > 3$ (see
Figure~\ref{fig:rp_period_age}).  However, Kepler discovered
$\approx$4{,}000 planets; and so unless there is a major discovery
bias, we would expect a larger fraction of the Kepler
planets---perhaps 10\%, or 400 planets---to be younger than 1\,Gyr.

\begin{figure}[!t]
	\begin{center}
		\leavevmode
		\includegraphics[width=0.49\textwidth]{f1.pdf}
	\end{center}
	\vspace{-0.6cm}
	\caption{
		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
    Planets younger than 1\,Gyr with ages more precise than a
    factor of three are emphasized. 
    ({\bf cite relevant papers?})
    Parameters are from the \citet{PSCompPars}.
		\label{fig:rp_period_age}
	}
\end{figure}

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{prot_teff_Santos19_Santos21_dquality.pdf}
      % vs prot_teff_Santos19_Santos21_clean0.pdf !
		}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{koi_mean_prot_teff_koi_X_S19S21dquality_keepgrazing.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Rotation periods for Kepler stars (left) and planets (right)}.
    The periods are from \citet{Santos_2019} and \citet{Santos_2021}
    (\citetalias{Santos_2019} and \citetalias{Santos_2021}), and the
    temperatures are primarily from \citet{Berger_2020a_catalog}.  The
    gray lines are ``mean fits'' from \citet{Bouma_2023} to the slow
    rotation sequences of open clusters.  The stellar sample shows
    only apparently single stars near the main sequence with $\log
    g$$>$4.2, RUWE$<$1.4, 3800$<$$T_{\rm eff}$/K$<$6200, that are not
    flagged as eclipsing binaries.  The planet sample has the same
    stellar cuts, and shows confirmed and candidate planets from the
    cumulative KOI table, requiring them to have a MES$>$10.
	}
	\label{fig:prot_vs_teff}
\end{figure*}

@@

>>>>

\author{et al}
%
%% Early experiments for Kepler rotation samples
%\author[0000-0001-7967-1795]{Elsa~K.~Palumbo}
%\affiliation{\caltech}
%
%% ?
%%\author{Lynne~A.~Hillenbrand}
%%\affiliation{\caltech}
%
%% HIRES observations and reductions
%\author[0000-0002-0531-1073]{H. Isaacson}
%\affiliation{\berkeley}
%% HIRES observations, reductions, JUMP infrastructure
%\author[0000-0001-8638-0320]{A. W. Howard}
%\affiliation{\caltech}
%%
%\author{HIRES observers to be named}
%\affiliation{Various}

@@

A limitation of the ``cluster-first'' approach is that only
$\approx$1\% of stars within the nearest few hundred parsecs can
currently be associated with their birth cluster
\citep[e.g.][]{Zari_2018,CantatGaudin_2020,Kounkel_2020,Kerr_2021}.
This means that there are many ``young'' (0.05-1\,Gyr) stars in the
field that are omitted by cluster-only searches.
Of order 40 transiting planets are currently known to be younger than
1\,Gyr, requiring an age precision $t/\sigma_t > 3$.
%TODO cite???
% (see
%Figure~\ref{fig:rp_period_age}).
However, Kepler discovered
$\approx$4{,}000 planets; and so unless there is a major discovery
bias, we would expect a larger fraction of the Kepler
planets---perhaps 10\%, or 400 planets---to be younger than 1\,Gyr.

%\begin{figure}[!t]
%	\begin{center}
%		\leavevmode
%		\includegraphics[width=0.49\textwidth]{f1.pdf}
%	\end{center}
%	\vspace{-0.6cm}
%	\caption{
%		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
%    Planets younger than 1\,Gyr with ages more precise than a
%    factor of three are emphasized. 
%    ({\bf cite relevant papers?})
%    Parameters are from the \citet{PSCompPars}.
%		\label{fig:rp_period_age}
%	}
%\end{figure}

@@

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{prot_teff_Santos19_Santos21_dquality.pdf}
		}
		\subfloat{
			\includegraphics[width=0.5\textwidth]{koi_mean_prot_teff_koi_X_S19S21dquality_keepgrazing.pdf}
		}
	\end{center}
	\vspace{-0.5cm}
	\caption{
    {\bf Rotation periods for Kepler stars (left) and planets (right)}.
    The periods are from \citet{Santos_2019} and \citet{Santos_2021}
    (\citetalias{Santos_2019} and \citetalias{Santos_2021}), and the
    temperatures are primarily from \citet{Berger_2020a_catalog}.  The
    gray lines are ``mean fits'' from \citet{Bouma_2023} to the slow
    rotation sequences of open clusters.  The stellar sample shows
    only apparently single stars near the main sequence with $\log
    g$$>$4.2, RUWE$<$1.4, 3800$<$$T_{\rm eff}$/K$<$6200, that are not
    flagged as eclipsing binaries.  The planet sample has the same
    stellar cuts, and shows confirmed and candidate planets from the
    cumulative KOI table, requiring them to have a MES$>$10.
	}
	\label{fig:prot_vs_teff}
\end{figure*}

@@

%\begin{figure*}[!t]
%	\begin{center}
%		\subfloat{
%			\includegraphics[width=0.5\textwidth]{rp_vs_period_scatter_20240415_colorbyage_strictyoung.pdf}
%		}
%		\subfloat{
%			\includegraphics[width=0.5\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_selsnrupper3_strictyoung.pdf}
%		}
%	
%		\subfloat{
%		\includegraphics[width=0.5\textwidth]{rp_vs_period_scatter_20240415_colorbyage_anyyoung.pdf}
%		}
%		\subfloat{
%			\includegraphics[width=0.5\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_anyyoung_anyyoung.pdf}
%		}
%	\end{center}
%	\vspace{-0.5cm}
%
%	\caption{
%		{\bf Radii, orbital periods, and ages of transiting exoplanets}.
%    Planets younger than 1\,Gyr are emphasized. 
%    Two possible selection functions are suggested...
%		\label{fig:rp_period_age_results}
%	}
%\end{figure*}

\begin{figure*}[!t]
	\begin{center}
		\subfloat{
			\includegraphics[width=0.8\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_selsnrupper3_strictyoung.pdf}
		}

		\subfloat{
			\includegraphics[width=0.8\textwidth]{rp_vs_period_scatter_20240415_colorbyage_showaux-gyro_anyyoung_anyyoung.pdf}
		}
	\end{center}
	\vspace{-0.5cm}

	\caption{
		{\bf Sizes, orbital periods, and rotation-based ages of young transiting exoplanets}.
		Two different selection functions for what defines a young, sub-gigayear, planet
		are imposed on the {\it top} and {\it bottom}: the former requires $t$>1\,Gyr with
		$t/\sigma_{\rm t}$>3, the latter also allows planets younger than 1\,Gyr at 2$\sigma$.
		Circles are previously reported measurements from the literature.
		Stars are newly calculated ages for Kepler planets from this work.
		Gray points are transiting planets from the NASA Exoplanet Archive older than
		1\,Gyr.
		The three age bins span 10-46\,Myr, 46-215\,Myr, and 215-1000\,Myr.
		\label{fig:rp_period_age_results}
	}
\end{figure*}
